Indexing Variation Graphs\footnote{This work was supported by the Wellcome Trust grant [098051].}

### 2016-04-20 paper/paper.tex
Added a simple CSA benchmark for the paper.

De~Bruijn graphs have several properties that make them useful for indexing purposes. Node keys are prefixes of the labels of all paths of length at least $k$ starting from the node. This makes it possible to sort the nodes unambiguosly by path labels. Every substring of the \kcollection{k} is the label of a path in the de~Bruijn graph, and every path label of length at most $k+1$ is a substring of the collection. In Section~\ref{sect:path-indexes}, we develop an index structure based on a generalization of de~Bruijn graphs.

@@

BWA-MEM \cite{Li2013} aligns reads to reference genomes by finding \emph{maximal exact matches} between the read and the reference. It stores both the reference and its reverse complement are stored in the same index, which allows it to extend the pattern in both directions. GCSA cannot use similar techniques, as we cannot guarantee that the length of the lexicographic range matching pattern $X$ is the same as the length of the range matching its reverse complement $\revcomp{X}$ (that the key set contains key $K$ if and only if it contains the reverse complement $\revcomp{K}$). Instead of using bidirectional BWT, we search for maximal exact matches using \LFmapping{} and the $\parent$ operation \cite{Ohlebusch2010a}. We have implemented a GCSA-based read aligner using that algoritm in vg \cite{Garrison2014-2016}.

@@

\begin{table}[t!]
\begin{center}
\caption{Performance benchmarks using an \orderk{64} GCSA2. Pattern length; number of patterns; number of matching patterns; number of matching nodes; number of distinct occurrences; average time for $[sp, ep] = \find(X)$, $\parent(sp, ep)$, and $\countq(sp, ep)$ queries in microseconds; and average time for $\locate(sp, ep)$ queries in microseconds or milliseconds.}\label{table:benchmark}
\begin{tabular}{c|cc|cc|cccc}
\hline
\textbf{Length} & \textbf{Patterns} & \textbf{Occs} & \textbf{Nodes} & \textbf{Values} & $\find()$ & $\parent()$ & $\countq()$ & $\locate()$ \\
\hline
16 & 351584 & 347453 & 1829M &  872M & 18.1 \textmu{}s & 0.40 \textmu{}s & 0.94 \textmu{}s & 26.2 ms \\
32 & 351555 & 333258 & 52.6M & 34.3M & 37.9 \textmu{}s & 0.28 \textmu{}s & 0.43 \textmu{}s & 468 \textmu{}s \\
64 & 351567 & 326101 & 0.33M & 1.35M &  142 \textmu{}s & 0.34 \textmu{}s & 0.57 \textmu{}s & 12.4 \textmu{}s \\
\hline
\end{tabular}
\end{center}
\end{table}

>>>>

De~Bruijn graphs have several properties that make them useful for indexing purposes. Node keys are prefixes of the labels of all paths of length at least $k$ starting from the node. This makes it possible to sort the nodes unambiguously by path labels. Every substring of the \kcollection{k} is the label of a path in the de~Bruijn graph, and every path label of length at most $k+1$ is a substring of the collection. In Section~\ref{sect:path-indexes}, we develop an index structure based on a generalization of de~Bruijn graphs.

@@

BWA-MEM \cite{Li2013} aligns reads to reference genomes by finding \emph{maximal exact matches} between the read and the reference. It stores both the reference and its reverse complement are stored in the same index, which allows it to extend the pattern in both directions. GCSA cannot use similar techniques, as we cannot guarantee that the length of the lexicographic range matching pattern $X$ is the same as the length of the range matching its reverse complement $\revcomp{X}$ (that the key set contains key $K$ if and only if it contains the reverse complement $\revcomp{K}$). Instead of using bidirectional BWT, we search for maximal exact matches using \LFmapping{} and the $\parent$ operation \cite{Ohlebusch2010a}. We have implemented a GCSA-based read aligner using that algorithm in vg \cite{Garrison2014-2016}.

@@

\begin{table}[t!]
\begin{center}
\caption{Performance benchmarks using an \orderk{64} GCSA2. Pattern length; number of patterns; number of matching patterns; number of matching nodes; number of distinct occurrences; average time for $[sp, ep] = \find(X)$, $\parent(sp, ep)$, and $\countq(sp, ep)$ queries in microseconds; and average time for $\locate(sp, ep)$ queries in microseconds or milliseconds.}\label{table:benchmark}
\begin{tabular}{c|cc|cc|cccc}
\hline
\textbf{Length} & \textbf{Patterns} & \textbf{Occs} & \textbf{Nodes} & \textbf{Values} & $\find()$ & $\parent()$ & $\countq()$ & $\locate()$ \\
\hline
16 & 351584 & 347453 & 1829M &  872M & 18.1 \textmu{}s & 0.40 \textmu{}s & 0.94 \textmu{}s & 26.2 ms \\
32 & 351555 & 333258 & 52.6M & 34.3M & 37.9 \textmu{}s & 0.28 \textmu{}s & 0.43 \textmu{}s & 468 \textmu{}s \\
64 & 351567 & 326101 & 0.33M & 1.35M & 88.9 \textmu{}s & 0.36 \textmu{}s & 0.42 \textmu{}s & 10.3 \textmu{}s \\
\hline
\end{tabular}
\end{center}
\end{table}
